---
title: "Modeltime"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Modeltime
Modeltime is a framework for building, evaluating, and deploying time series models in R. It provides a consistent interface for working with various time series models, making it easier to compare and select the best model for your data.

# Power of modeltime
Modeltime allows you to:
- Build and evaluate multiple time series models using a consistent interface.
- Compare model performance using various metrics.
- Easily deploy models for forecasting.
- Integrate with other R packages for data manipulation and visualization.

# Steps in modeltime
1. **Data Preparation**: Prepare your time series data, ensuring it is in the correct format.
2. **Model Specification**: Specify the models you want to use for forecasting.
3. **Model Fitting**: Fit the models to your training data.
4. **Model Evaluation**: Evaluate the models using various metrics to determine their performance.
5. **Model Forecasting**: Use the best-performing model to make forecasts on new data.
6. **Model Deployment**: Deploy the model for production use, allowing for real-time forecasting.

packages needed:
```{}
library(modeltime)
library(tidymodels)
library(timetk)
library(dplyr)
library(lubridate)
library(ggplot2)
library(zoo)
library(modeltime.ensemble)

```

# Example of modeltime usage

## Load necessary libraries
```{r}
library(modeltime)
library(tidymodels)
library(timetk)
library(dplyr)
library(lubridate)
library(ggplot2)
library(zoo)
library(modeltime.ensemble)

```

## Load example data

```{r}
data <- tibble(
  date  = as.Date(as.yearmon(time(AirPassengers))),
  value = as.numeric(AirPassengers)
)
```

the data should be in a tibble format with a date column and a value column.

## plotling the data
```{}
data %>% plot_time_series(date, value)
```

this will plot the time series data, allowing you to visualize trends and patterns over time. 

## Split the data into training and testing sets

```{r}
splits <- initial_time_split(data, prop = 0.9)
```

## Specify the model

```{r}
model_arima <- arima_reg() %>%
  set_engine("auto_arima") %>%
  fit(value ~ date, data = training(splits))

model_ets <- exp_smoothing(seasonal_period = 12) %>%
  set_engine("ets") %>%
  fit(value ~ date, data = training(splits))

model_prophet <- prophet_reg() %>%
  set_engine("prophet") %>%
  fit(value ~ date, data = training(splits))
```

## Create a modeltime table
```{r}
models_tbl <- modeltime_table(
  model_arima,
  model_ets,
  model_prophet
)

```
model table is a table that contains all the models you want to compare.

## to know the summary of the models

```{r}
# ARIMA Summary
models_tbl %>% 
  pluck_modeltime_model(1)

# ETS Summary
models_tbl %>% 
  pluck_modeltime_model(2)

# Prophet Model (no summary method, inspect structure)
models_tbl %>% 
  pluck_modeltime_model(3)

```

## Evaluate the models on the testing set

```{r}
calibration_tbl <- models_tbl %>%
  modeltime_calibrate(new_data = testing(splits))
```

this will give you a table with the calibrated models.

```{r}
calibration_tbl %>%
  modeltime_forecast(
    new_data = testing(splits),
    actual_data = data
  )
```
this will give you the forecasted values for the testing set.

## Plot the forecasted values
```{}
calibration_tbl %>%
  modeltime_forecast(
    new_data = testing(splits),
    actual_data = data
  ) %>%
  plot_modeltime_forecast()
```
This will plot the forecasted values along with the actual values from the testing set.

# Evolution of models by metrics
```{r}
calibration_tbl %>%
  modeltime_accuracy() %>%
  arrange(rmse)
```
This will give you a table with the accuracy metrics for each model, sorted by RMSE (Root Mean Square Error).

## Refitting 

```{r}
# Refit for the entire dataset
refit_tbl <- calibration_tbl %>%
  modeltime_refit(data = data)
```
This will refit the models using the entire dataset, allowing you to use them for forecasting on new data.

## Forecasting on new data
```{r}
# Forecasting for the next 12 months
future_tbl <- data %>%
  future_frame(.date_var = date, .length_out = 12)
```
mention the future frame function which creates a future frame for forecasting.

## Forcasting
```{r}
forecast_future <- refit_tbl %>%
  modeltime_forecast(new_data = future_tbl, actual_data = data)
```
This will generate forecasts for the next 12 months based on the refitted models.

## Plot the future forecast
```{}
forecast_future %>%
  plot_modeltime_forecast()
```
This will plot the forecasted values for the next 12 months, allowing you to visualize the expected trends and patterns in the data.

# Viewing the forecasted values
```{r}
forecast_future %>%
  select(.model_desc, .index, .value, .conf_lo, .conf_hi) %>% head()

forecast_future %>% head()
```

This will display the forecasted values along with their confidence intervals, allowing you to see the expected range of values for the next 12 months.

modeltime also allow the user to create ensembles of models, which can improve forecasting accuracy by combining the strengths of multiple models.
which means hybrid models can be created using modeltime.ensemble package.

## 📊 Summary of Ensemble Methods in Modeltime

This allows advanced ensembling with meta-learning or stacking, but it requires manual implementation or integration with workflowsets or stacks.

| Function            | Type                  | Weights Required | Key Use Case                             |
|---------------------|------------------------|------------------|------------------------------------------|
| `ensemble_average()`  | Simple Average         | ❌               | Quick hybrid model                       |
| `ensemble_weighted()` | Weighted Average       | ✅               | Give more importance to better models    |
| Custom Stacking       | Meta-Learning / Blend  | ✅               | Advanced ML ensemble (manual)            |

# Example of Ensemble specification
```{}
ensemble_model <- ensemble_average(models_tbl)
ensemble_model <- ensemble_weighted(models_tbl, weights = c(0.6, 0.3, 0.1))

```
by these functions, you can create ensemble models that combine the predictions of multiple individual models, allowing for more robust forecasting.

# example of Hybrid models

## data
```{r}
data("AirPassengers")
air_df <- tibble(
  date  = as.Date(as.yearmon(time(AirPassengers))),
  value = as.numeric(AirPassengers)
)

```

## Split the data into training and testing sets

```{r}
splits <- initial_time_split(air_df, prop = 0.9)

```
## Specify the models
```{r}
model_fit_arima <- arima_reg() %>%
  set_engine("auto_arima") %>%
  fit(value ~ date, data = training(splits))

model_fit_ets <- exp_smoothing() %>%
  set_engine("ets") %>%
  fit(value ~ date, data = training(splits))

model_fit_prophet <- prophet_reg() %>%
  set_engine("prophet") %>%
  fit(value ~ date, data = training(splits))
```

## Create a modeltime table
```{r}
models_tbl <- modeltime_table(
  model_fit_arima,
  model_fit_ets,
  model_fit_prophet
)
```
# 🌟 Create an Ensemble Model (Average)
```{r}
ensemble_model <- ensemble_average(models_tbl)
ensemble_tbl <- modeltime_table(ensemble_model)

```
# Combine all 4 models
```{r}
all_models_tbl <- bind_rows(models_tbl, ensemble_tbl)
```

```{r}
all_models_tbl %>% pluck_modeltime_model(4)
```




# Evaluate the models on the testing set
```{r}
calibration_tbl <- all_models_tbl %>%
  modeltime_calibrate(new_data = testing(splits))
```

# Show Accuracy Metrics

```{r}
calibration_tbl %>%
  modeltime_accuracy() %>%
  arrange(rmse)
```
# Plot Predictions vs Actuals (Testing Period)
```{}
calibration_tbl %>%
  modeltime_forecast(
    new_data = testing(splits),
    actual_data = air_df
  ) %>%
  plot_modeltime_forecast()
```

# Refitting for the entire data
```{r}
all_models_refit <- all_models_tbl %>%
  modeltime_refit(data = air_df)
```

# Forecasting for the next 12 months
```{r}
future_tbl <- air_df %>%
  future_frame(.date_var = date, .length_out = 12)

forecast_tbl <- all_models_refit %>%
  modeltime_forecast(new_data = future_tbl, actual_data = air_df)

```
# Plot Future Forecast
```{}
forecast_tbl %>%
  plot_modeltime_forecast()
```
# View Forecasted Values
```{r}
forecast_tbl %>%
  head()
```

