---
title: "lubridate"
output: github_document
---

lubridate is an R package that makes it easier to work with dates and times. It simplifies parsing, extraction, modification, and arithmetic operations on dates.

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(lubridate)
```

# 1. Parsing Dates and Date-Times

Common lubridate Parsing Functions:

-   ymd() – Year-Month-Day

-   dmy() – Day-Month-Year

-   mdy() – Month-Day-Year

-   ydm(), myd(), dym() – Other formats

-   ymd_hms(), dmy_hms() – Include time (hour, minute, second)

```{r echo=TRUE, message=FALSE, warning=FALSE}
ymd("2025-07-17")             # "2025-07-17"
dmy("17-07-2025")             # "2025-07-17"
mdy("July 17, 2025")          # "2025-07-17"
ydm("2025-17-07")             # "2025-07-17"
ymd_hms("2025-07-17 14:30:00")# "2025-07-17 14:30:00 UTC"

```

# 2. Extracting Date-Time Components Functions:

-   year(), month(), day()

-   hour(), minute(), second()

-   wday(), yday(), mday()

-   week(), quarter(), semester()

```{r echo=TRUE, message=FALSE, warning=FALSE}
dt <- ymd_hms("2025-07-17 14:45:30")

year(dt)      # 2025
month(dt)     # 7
month(dt, label = TRUE)        # Jul
month(dt, label = TRUE, abbr = FALSE) # July
day(dt)       # 17
hour(dt)      # 14
minute(dt)    # 45
second(dt)    # 30
wday(dt, label = TRUE)         # Thu
wday(dt, label = TRUE, abbr = FALSE) # Thursday
yday(dt)      # day of the year ( 1 to 365)
quarter(dt)   # quarter of the year (1 to 4)
week(dt)     # week of the year (1 to 53)
semester(dt)  # semester of the year (1 or 2)
mday(dt)      # day of the month (17)
```


# 4. Time Zones ✅ Functions:

-   tz() – Get time zone

-   with_tz() – Convert to new zone (adjusts time)

-   force_tz() – Change zone without changing clock time

```{r echo=TRUE, message=FALSE, warning=FALSE}
dt <- ymd_hms("2025-07-17 10:00:00", tz = "Asia/Kolkata")

with_tz(dt, "UTC")     # "2025-07-17 04:30:00 UTC"
force_tz(dt, "UTC")    # "2025-07-17 10:00:00 UTC"

```

# 5. Rounding Dates

✅ Functions:

-   round_date()

-   floor_date()

-   ceiling_date()

```{r echo=TRUE, message=FALSE, warning=FALSE}
x <- ymd_hms("2025-07-17 10:34:50")

round_date(x, "hour")    # "2025-07-17 11:00:00 UTC"
floor_date(x, "hour")    # "2025-07-17 10:00:00 UTC"
ceiling_date(x, "hour")  # "2025-07-17 11:00:00 UTC"

```

# 6. Time Arithmetic

-   +, - – Add/subtract durations

-   days(), months(), years()

-   duration(), period()

```{r echo=TRUE, message=FALSE, warning=FALSE}
date <- ymd("2025-07-17")

date + days(10)               # "2025-07-27"
date + period(1, "month")     # "2025-08-17"
date + duration(3600)         # "2025-07-17 01:00:00 UTC"

```

# 7. Intervals and Durations

✅ Functions:

-   interval(start, end)

-   %within% – Check if a date falls within an interval
```{r echo=TRUE, message=FALSE, warning=FALSE}
int <- interval(ymd("2025-07-01"), ymd("2025-07-31"))

# length in days
int_length(int)/86400

target <- ymd("2025-07-17")
target %within% int

```



#  8. Other Useful Functions

Below are some handy functions in `lubridate` for working with dates and times:

| **Function**       | **Purpose**                               |
|--------------------|-------------------------------------------|
| `today()`          | Current date                              |
| `now()`            | Current date-time                         |
| `leap_year()`      | Check for leap year                       |
| `is.Date()`        | Check if object is of Date class          |
| `make_date()`      | Create a date from year, month, and day   |
| `make_datetime()`  | Create a datetime from components         |

```{r echo=TRUE, message=FALSE, warning=FALSE}
today()
now()
leap_year(2024)
make_date(2025, 12, 31)
make_datetime(2025, 12, 31, 14, 30, 0)

```

#  Summary Table of `lubridate` Functions

| **Category**      | **Functions**                            |
|-------------------|-------------------------------------------|
| Parsing           | `ymd()`, `dmy()`, `mdy()`, `ymd_hms()`    |
| Extraction        | `year()`, `month()`, `day()`, `hour()`    |
| Modification      | `year() <-`, `month() <-`, etc.           |
| Time Zones        | `with_tz()`, `force_tz()`, `tz()`         |
| Rounding          | `round_date()`, `floor_date()`            |
| Arithmetic        | `days()`, `months()`, `duration()`        |
| Intervals         | `interval()`, `%within%`                  |
| Current Dates     | `today()`, `now()`, `leap_year()`         |


# System date functions


| **Function**   | **Returns**          | **Class**   | **Time Included** |
|----------------|----------------------|-------------|--------------------|
| `Sys.Date()`   | System Date          | `Date`      | ❌ No              |
| `Sys.time()`   | System Date-Time     | `POSIXct`   | ✅ Yes             |
| `now()`        | lubridate Date-Time  | `POSIXct`   | ✅ Yes             |

```{r}
Sys.Date()          # "2025-07-17"
Sys.time()          

```

## Formate changing 
```{r}
format(Sys.time(), "%Y-%m-%d")   # Just the date
format(Sys.time(), "%H:%M:%S")   # Just the time
format(Sys.time(), "%Y-%m-%d %H:%M:%S") # Full date-time  

```
## Adding 
```{r}
Sys.Date()+ 1  # Adds one day
Sys.Date()+ month(1)  # Adds one month
Sys.Date() + years(1)  # Adds one year
Sys.Date() + days(10)  # Adds 10 days

```

```{r}
Sys.time() + hours(2)  # Adds 2 hours
Sys.time() + minutes(30)  # Adds 30 minutes
Sys.time() + seconds(45)  # Adds 45 seconds
Sys.time() + days(1)  # Adds 1 day
Sys.time() + weeks(1)  # Adds 1 week
Sys.time() + months(1)  # Adds 1 month
Sys.time() + years(1)  # Adds 1 year
```

